---
title: 'Reproducible Research: Peer Assessment 1'
author: "GK"
output:
    html_document:
    keep_md: yes
---

========================================

## Loading and preprocessing the data
```{r load_data,echo=TRUE}
unzip(zipfile="activity.zip")
data<- read.csv("activity.csv")
```

## What is mean total number of steps taken per day? 
```{r histogram,echo=TRUE}
library(ggplot2)
total.steps<-data.frame(tapply(data$steps,data$date,sum,na.rm=T))
total.steps$date<- rownames(total.steps)
colnames(total.steps)<- c("Steps","Date")
ggplot(total.steps,aes(x=Steps)) + geom_histogram(fill="steelblue",binwidth = 1000) + labs(title="Histogram of the total number of steps taken each day",x="Steps",y="Frequency")
```

```{r mean_median,echo=TRUE}
mea<- mean(total.steps$Steps,na.rm=T)
med<- median(total.steps$Steps,na.rm=T)
```
Mean : `r mea`
Median : `r med`

## What is the average daily activity pattern?
```{r time_plot,echo=TRUE}
library(ggplot2)
steps.interval<- data.frame(tapply(data$steps,data$interval,mean,na.rm=T))
steps.interval$interval<- rownames(steps.interval)
colnames(steps.interval)<- c("Steps","Interval")
ggplot(steps.interval,aes(x=as.numeric(Interval),y=Steps)) + geom_line(color="red") + labs(title="Time series plot of the average number of steps taken",x="5 Minute Intervals",y="Steps")
```

```{r max_interval,echo=TRUE}
maxInterval<- steps.interval[which.max(steps.interval$Steps),]
```
The 5-minute interval that, on average, contains the maximum number of steps : `r maxInterval`

## Imputing missing values
```{r how_many_missing,echo=TRUE}
missing<- is.na(data$steps)
tab<-table(missing)
```
`r tab`

All of the missing values are filled in with mean value for that 5-minute
interval.

```{r fill_na,,echo=TRUE}
fill.data<- function(steps,interval){
    filled<-NA
    if(!is.na(steps))
        filled<-c(steps)
    else
        filled<- (steps.interval[steps.interval$Interval==interval,"Steps"])
    return(filled)
}
filled.data<- data
filled.data$steps<- mapply(fill.data,filled.data$steps,filled.data$interval)
```

Histogram of the total number of steps taken each day after missing values are imputed
```{r new_data_hist,echo=TRUE}
new.total.steps<- data.frame(tapply(filled.data$steps,filled.data$date,sum,na.rm=T))
new.total.steps$date<-rownames(new.total.steps)
colnames(new.total.steps)<- c("Steps","Date")
ggplot(new.total.steps,aes(x=Steps)) + geom_histogram(fill="green",binwidth = 1000) + labs(x="Steps",y="Frequency",title="Histogram")
```

Calculate mean and median of the new modified data.
```{r new_summary,echo=TRUE}
mea1<- mean(new.total.steps$Steps)
med1<- median(new.total.steps$Steps)
```

Old Mean : `r mea`
Old Median : `r med`


New Mean : `r mea1`
New Median : `r med1`

## Are there differences in activity patterns between weekdays and weekends?
```{r weekdata,echo=TRUE}
filled.data$datetype<- ifelse(as.POSIXlt(filled.data$date)$wday %in% c(0,6),"weekend","weekday")
```

Now, let's make a panel plot containing plots of average number of steps taken
on weekdays and weekends.
```{r panel_plot,echo=TRUE}
library(dplyr)
day.data<- filled.data %>% group_by(datetype,interval) %>% summarize(totalSteps=mean(steps))
ggplot(day.data,aes(x=interval,y=totalSteps,color=datetype)) + geom_line() + facet_grid(datetype~.) + labs(x="5 Minute Interval",y="Steps",title="Average steps taken in Weekdays/Weekend")
```
